### 2014-05-20 Paper1/paper1.tex
Pushed and saved

\section*{Introduction}
As machines interact more and more with humans, there is a growing need for a rigorous way to understand what people are trying to do, even if they become disturbed, distracted or influenced by pathology. One can assume that a person who is familiar with a particular task will try to do the same thing each time provided all other conditions are equal. This presumes that what people do is on average what they intended to do and that their goal remains unchanged. However, if there is a disturbance, it is unclear what happens to a person's intentions because action is decoupled from intention. Lacking a means to discern that intention and changes in intention leads to challenges in human machine interactions. Hence a comprehensive method for understanding what someone means to do (rather than what they actually do or claim to have wanted to do) as they interact within uncertain environments will prove useful in any area where a person might be thwarted from their intent such as national security, operation of machinery, athletic and music performance, rehabilitation, human augmentation, and artificial intelligence.

@@

Understanding the intent has many benefits, and one may be a better way to understand the composition of discrete movements into their subunits. Recent research has strived to understand the composition of movements intention during both infant development\cite{von1979observations} and stroke recovery\cite{rohrer2004submovements}, distinct straight-line subunits with bell-shaped speed profiles increasingly overlap and blend until, as in healthy adult reaching\cite{woodworth1899accuracy}, they become indistinguishable. Flash and Henis \cite{flash1991arm} demonstrated that despite this apparent blending, cued corrections can only be produced at predictable places within the reach. They also showed that these modifications appear to take place by superimposing additional straight-line subunits onto the original movement. Novak et al. \cite{novak2002use} detected corrective subunits following disturbance during knob turning. These studies show planning and correction in discrete subunits, but did not analyze the actual intent and hence cannot be compared to a known, correct trajectory except in the trivial case that movement is undisturbed. A disturbed trajectory's intention is hidden without proper inversion of the model plan for movement. Here we develop a method to extract intent and gain a clearer understanding of subunits from disturbed motions.

@@

\subsection*{Generation of synthetic data using simple dynamic simulation}
We choose one prevailing arm model, that of Burdet et al. \cite{burdet2006stability}, along with the anatomical mass distribution and geometry parameters of Winter \cite{winter2009biomechanics} to create a model for estimating intent. The model of Burdet et al \cite{burdet2006stability} was adapted to reflect measured physical parameters of the subject who provided the reach. Segment lengths were measured directly while gross body mass was self-reported. These values were converted to segment masses, centers of gravity, and moments of inertia inertia using anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics}. The muscle torque used to calculate stiffness was approximated by looking backward in time a small amount (500 microseconds). Desired trajectory in time was simulated as a minimum jerk, 5th order polynomial of duration 700ms starting and ending with zero velocity and acceleration.
Simulated reaches were 15 cm long, beginning 38 cm out from and 5.7 cm left of the right shoulder then proceeding right.  We used the following types of perturbing forces for each combination of distance and direction: 
Pulse forces applied in one of the the two directions perpendicular to the direction of movement began when the subject had moved either 10\% or 50\% of the distance to the target and lasted for 150ms.
Noise forces began once the subject moved 3 mm, and lasted for the duration of the motion. The forces were drawn from a white noise generator drawn at 1000 Hz with flat power spectral density of 1N, and then passed through a 4th order low-pass Butterworth filter with cutoff $10\pi$ rad/s.

@@

\subsection*{Human Subjects}
The human data trajectories analyzed here are drawn from a four subjects who gave informed consent in accordance with Northwestern University Institutional Review Board. Two male and two female right-handed subjects (ages 24 to 30) performed the reaches with their right arm and was not compensated. Subjects' arm segment lengths were directly measured \textit{in situ} while body mass was self-reported.

\subsection*{Apparatus}
A planar manipulandum (described in Patton and Mussa-Ivaldi \cite{patton2004robot}) was programmed to compensate and minimize any handle friction or mass. The MATLAB XPC-TARGET package (Natick, MA) was used to render this force environment at 1000 Hz and data were collected at 1000 Hz.  Visual feedback was performed at 60 Hz using OpenGL. Closed loop data transmission time (position measurement to completed rendering to recognition of rendering by the position measurement system) was <8 ms, ensuring a visual delay less than one 60 Hz frame.

@@

\subsection*{Model Viability}
Sensitivity analysis fortunately revealed that variance in intended trajectories due to estimated uncertainty in model parameters is lower than the natural variation in undisturbed motions.  Simulated point-to-point reaches disturbed by either filtered white noise forces or a pulse force perpendicular to the direction of the reach (figure 1) were extracted in the presence of 220,000 variations upon the parameters (table 1) according to the methods of Saltelli et al. \cite{saltelli2010variance}. Expected variation due to parameter uncertainty was 0.86 microns for pulse forces and 1.5 microns for filtered white forces. Variation in recorded undisturbed point-to-point reaching under the same time and reach distance conditions was 2.93 microns. These variations both fall beneath the instantaneous accuracy of the robot recording the motion (tenths of a millimeter). In other words, while some inversion processes might be highly sensitive to model uncertainty, our process for recovering intent from action is much more precise than its input.

@@

\subsection*{Human Intent Trajectories}
Extraction of intent from human point-to-point reaching revealed straight-line movement that persisted for hundreds of milliseconds after the onset of disturbing forces (figure 3). In both pulse timing conditions, intent remained straight line towards the target, supporting the approach.  When disturbances occurred early in reaching, intent showed some signs of correction mid-reach. When disturbances occurred late in reaching, intent tended to continue to onto the target with change of intent occurring on or near the target (figure 3 lower plots). 
Patterns across all trials and subject revealed some systematic tendencies. For approximately 200 milliseconds after the onset of disturbance (figure 4 compare individual points). At this point intent significantly deviated from the baseline region of tendency. As the movement completed, the hand position and intent re-converged and become statistically indistinguishable at approximately 400ms after the onset of disturbance (figure 4 shaded areas).

\section*{Discussion}
We sought to determine the plausibility and robustness of an algorithm that determines a person’s intent of a motion, even if there disturbances. Sensitivity analysis on synthetic data revealed that the method was plausible and response to parameter variations was more robust than human reaching, but significant misestimation could occur if the stiffness was inaccurately modeled. Examination of intended trajectories from force pulse-disturbed human reaching confirmed that these intentions remained statistically indistinguishable from undisturbed reaching for ~100ms following the onset of disturbance. This test provides a method for testing and potentially falsifying our extraction. Differences after this point could be attributed to many factors including unmodeled reflexes, changes in arm stiffness, or genuine change in movement intent. In the context of the sensitivity analysis, it appears most likely that this primarily represents a change in intent and thus a first look at how intent changes when faced with a force disturbance.

>>>>

\section*{Introduction}
As machines interact more and more with humans, there is a growing need for a rigorous way to understand what people are trying to do, even if they become disturbed, distracted, or influenced by pathology. One can assume that a person who is familiar with a particular task will try to do the same thing each time provided all other conditions are equal. This presumes that what people do is on average what they intended to do and that their goal remains unchanged. However, if there is a disturbance, it is unclear what happens to a person's intentions because action is decoupled from intention. Lacking a means to discern that intention and changes in intention leads to challenges in human-machine interactions. Hence a comprehensive method for understanding what someone means to do (rather than what they actually do or claim to have wanted to do) as they interact within uncertain environments will prove useful in any area where a person might be thwarted from their intent such as national security, operation of machinery, athletic and music performance, rehabilitation, human augmentation, and artificial intelligence.

@@

Examination of movement intent during disturbance should allow us to test whether or not that intent is elemental or composed of subunits. Recent research has strived to understand the composition of movement intention during both infant development\cite{von1979observations} and stroke recovery\cite{rohrer2004submovements} where distinct straight-line subunits with bell-shaped speed profiles increasingly overlap and blend until, as in healthy adult reaching\cite{woodworth1899accuracy}, they become indistinguishable. Past attempts to isolate subunits from healthy adult reaching are inconclusive. Flash and Henis \cite{flash1991arm} demonstrated that  cued corrections produce recoverable subunits with a predictable shape and structure, but Liu and Todorov were able to reproduce this without using subunits \cite{liu2007evidence}. Novak et al. \cite{novak2002use} detected corrective subunits following disturbance during knob turning, but lacking access to the intent they instead relied on kinematic features of the disturbed trajectory. Here we develop a method to extract intent from disturbed movement, which will allow us to observe how intent changes during uncued replanning and whether that change is discrete or continuous.

@@

\subsection*{Generation of synthetic data using simple dynamic simulation}
We choose one prevailing arm model, that of Burdet et al. \cite{burdet2006stability}, along with the anatomical mass distribution and geometry parameters of Winter \cite{winter2009biomechanics} to create a model for estimating intent. The model of Burdet et al \cite{burdet2006stability} was adapted to reflect measured physical parameters of the subject who provided the reach. Segment lengths were measured directly while gross body mass was self-reported. These values were converted to segment masses, centers of gravity, and moments of inertia using anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics}. The muscle torque used to calculate stiffness was approximated by looking backward in time a small amount (500 microseconds). Desired trajectory in time was simulated as a minimum jerk, 5th order polynomial of duration 700ms starting and ending with zero velocity and acceleration.
Simulated reaches were 15 cm long, beginning 38 cm out from and 5.7 cm left of the right shoulder then proceeding right.  We used the following types of perturbing forces for each combination of distance and direction: 
Pulse forces applied in one of the the two directions perpendicular to the direction of movement began when the subject had moved either 10\% or 50\% of the distance to the target and lasted for 150ms.
Noise forces began once the subject moved 3 mm, and lasted for the duration of the motion. The forces were drawn from a white noise generator at 1000 Hz with flat power spectral density of 1N, and then passed through a 4th order low-pass Butterworth filter with cutoff $10\pi$ rad/s.

@@

\subsection*{Human Subjects}
The human data trajectories analyzed here are drawn from a four subjects who gave informed consent in accordance with Northwestern University Institutional Review Board. Two male and two female right-handed subjects (ages 24 to 30) performed the reaches with their right arm and were not compensated. Subjects' arm segment lengths were directly measured \textit{in situ} while body mass was self-reported.

\subsection*{Apparatus}
A planar manipulandum (described in Patton and Mussa-Ivaldi \cite{patton2004robot}) was programmed to compensate and minimize any handle friction or mass. The MATLAB XPC-TARGET package (Natick, MA) was used to render this force environment at 1000 Hz and data were collected at 1000 Hz.  Visual feedback was performed at 60 Hz using OpenGL. Closed loop data transmission time (position measurement to completed rendering to recognition of rendering by the position measurement system) was less than 8 ms, ensuring a visual delay less than one 60 Hz frame.

@@

\subsection*{Model Robustness}
Sensitivity analysis fortunately revealed that variance in intended trajectories due to estimated uncertainty in model parameters is lower than the natural variation in undisturbed motions.  Simulated point-to-point reaches disturbed by either filtered white noise forces or a pulse force perpendicular to the direction of the reach (figure 1) were extracted in the presence of 220,000 variations upon the parameters (table 1) according to the methods of Saltelli et al. \cite{saltelli2010variance}. Expected variance due to parameter uncertainty was $0.86  \mu m^2$ for pulse forces and $1.5 \mu m^2$ for filtered white forces. Variance in recorded undisturbed point-to-point reaching under the same time and reach distance conditions was $2.93 \mu m^2$. These variances all fall beneath the instantaneous accuracy of the robot recording the motion (tenths of a millimeter). In other words, while some inversion processes might be highly sensitive to model uncertainty, our process for recovering intent from action is much more precise than its input.

@@

\subsection*{Human Intent Trajectories}
Extraction of intent from human point-to-point reaching revealed straight-line movement that persisted for hundreds of milliseconds after the onset of disturbing forces (figure 3). In both pulse timing conditions, intent remained straight line towards the target, supporting the approach.  When disturbances occurred early in reaching, intent showed some signs of correction mid-reach. When disturbances occurred late in reaching, intent tended to continue onto the target with intent changing on or near the target (figure 3 lower plots). Patterns across all disturbance timings and subjects revealed some systematic tendencies. Force-disturbed hand paths deviate from undisturbed reaching approximately 25 milliseconds after the onset of disturbance while intention remains comparable to undisturbed reaching for 100 milliseconds. While the hand returns within 500 milliseconds to undisturbed levels of error, intention is altered for an entire second (figure 4 shaded areas).

\section*{Discussion}
We sought to determine the plausibility and robustness of an algorithm that determines a person's intent during reaching motion, even if there disturbances. Sensitivity analysis on synthetic data revealed that the method was plausible. Response to parameter variations was more robust than human reaching, but significant misestimation could occur if the stiffness was inaccurately modelled. Examination of intended trajectories from force pulse-disturbed human reaching confirmed that these intentions remained statistically indistinguishable from undisturbed reaching for about 100ms following the onset of disturbance. This test provides a method for testing and potentially falsifying our extraction. Differences after this point could be attributed to many factors including unmodeled reflexes, changes in arm stiffness, or genuine change in movement intent. In the context of the sensitivity analysis, it appears most likely that this primarily represents a change in intent and thus a first look at how intent changes when faced with a force disturbance.
